What's in a Name? \\ Linear Temporal Logic \\ Literally Represents Time Lines

### 2023-08-29 paper/main.tex
minor change to timeline length definition
minor change
added small fix
clarification

%KYR: FIX THIS! We need to make it clear the relationship between Algorithms 1..4 and the figure displaying the high-level algorithm.
The algorithm is outlined below (see Algorithms 1 - 4). %\cref{alg:ba2wregex}

@@

\begin{definition}[Timeline Length]
    We define the timeline length of a regular expression $\tllen(A)$ recursively as follows.
    \begin{align*}
        \tllen(\emptyset)   &= -\infty\\  %KYR:FIX!!!
        \tllen(\epsilon)        &= 0\\
        \tllen(a)           &= 1\\
        \tllen(A_1 A_2)     &= \tllen(A_1) + \tllen(A_2)\\
        \tllen(A_1 + A_2)   &= \max(\tllen(A_1), \tllen(A_2))\\
        \tllen(A^*)         &= \tllen(A)
    \end{align*}
    We also extend this definition to $\omega$-regular expressions as follows.
    \begin{align*}
        \tllen(A^\omega)    &= \tllen(A)\\
        \tllen(AB)          &= \tllen(A) + \tllen(B)\\
        \tllen(B_1 + B_2)   &= \max(\tllen(B_1), \tllen(B_2))
    \end{align*}
    Then $\tllen$ captures the length of the longest path in our timeline graph visualization.
\end{definition}

@@

%KYR: Here we nicely show (with pictures as much as possible) all of the individual statistics and results we gathered (note aggregation occurs below).
%\paragraph{Real-world datasets}
{\bf Feasibility.}
We study the feasibility of our tool on a benchmark suite of LTL formulas gathered from real world use cases.
%
We gather a test suite totalling 91 formulas from two real-world requirement specification applications. We take $6$ specifications from NASA's Automated Airspace Concept (AAC)~\cite{GCMTR16}. We also collect formulas from the Acacia suite of examples~\cite{RV10}. We reap $85$ formulas from the suite of $23$ examples, by generating separate visualizations for each line of specification. We choose to visualize each formula individually, rather than visualize a complete set of (conjuncted) specifications, as a complete specification set may describe a large, complex system, and we only target the use case of validating individual formulas. % We choose to visualize each formula specifically, rather than complete specifications, as

@@

%\paragraph{Scalability}
{\bf Scalability.}
%KYR: FIX!!! The description of the exact experiment run is not clear and seems to contradict the points in the above paragraph...
We measure the scalability of our tool by running it against a set of \emph{scalable} LTL formulas. We define a scalable LTL formula as one that describes an $n$-bit number using a binary counter, as generated by~\cite{RV10}. As $n$ gets larger, the LTL formula required to encode the $n$-bit counter scales in size. \cref{fig:scalability} shows that the timeline length grows approximately exponentially with the number of bits in the counter, which matches the result in Rozier and Vardi~\cite{RV10}.

>>>>

%KYR: FIX THIS! We need to make it clear the relationship between Algorithms 1..4 and the figure displaying the high-level algorithm.
The algorithm to convert \Buchi automata to $\omega$-regular expressions \texttt{ba2wregex} combines Algorthms 1--4 below. %\cref{alg:ba2wregex}

@@

\begin{definition}[Timeline Length]
    We define the timeline length of a regular expression $\tllen(A)$ recursively as follows.
    \begin{align*}
        \tllen(\emptyset)   & \quad \text{undefined} \\
        \tllen(\epsilon)    &= 0\\
        \tllen(a)           &= 1\\
        \tllen(A_1 A_2)     &= \tllen(A_1) + \tllen(A_2)\\
        \tllen(A_1 + A_2)   &= \max(\tllen(A_1), \tllen(A_2))\\
        \tllen(A^*)         &= \tllen(A)
    \end{align*}
    We also extend this definition to $\omega$-regular expressions as follows.
    \begin{align*}
        \tllen(A^\omega)    &= \tllen(A)\\
        \tllen(AB)          &= \tllen(A) + \tllen(B)\\
        \tllen(B_1 + B_2)   &= \max(\tllen(B_1), \tllen(B_2))
    \end{align*}
    Then $\tllen$ captures the length of the longest path in our timeline graph visualization.
\end{definition}

@@

%KYR: Here we nicely show (with pictures as much as possible) all of the individual statistics and results we gathered (note aggregation occurs below).
%\paragraph{Real-world datasets}
{\bf Feasibility.}
We study the feasibility of our tool on a benchmark suite of LTL formulas gathered from real world use cases.
%
We gather a test suite totalling 91 formulas from two real-world requirement specification applications. We take $6$ specifications from NASA's Automated Airspace Concept (AAC)~\cite{GCMTR16}. We also collect formulas from the Acacia suite of examples~\cite{RV10}: $85$ formulas from the suite of $23$ collections of formulas were collected. We choose to visualize each formula individually, rather than visualize a complete set of (conjuncted) specifications, as a complete specification set may describe a large, complex system, and we only target the use case of validating individual formulas. The set of formulas being tested can be found \href{https://github.com/EULIR/ltl-explainability/blob/main/ltl-formulas/benchmark-ltls.txt}{here}.

@@

%\paragraph{Scalability}
{\bf Scalability.}
%KYR: FIX!!! The description of the exact experiment run is not clear and seems to contradict the points in the above paragraph...
We measure the scalability of our tool by running it against a set of \emph{scalable} LTL formulas. We define a scalable LTL formula as one that describes an $n$-bit number using a binary counter, as generated by~\cite{RV10}. As $n$ gets larger, the LTL formula required to encode the $n$-bit counter scales in size. \cref{fig:scalability} shows that the timeline length grows approximately exponentially with the number of bits in the counter, which matches the result in Rozier and Vardi~\cite{RV10}. The set of formulas being tested can be found \href{https://github.com/EULIR/ltl-explainability/tree/main/ltl-formulas/binary_counter}{here}.
